### 2016-06-08 capability-machine.tex
Changed some of the definition for the value relation.
Updated the value interpretation definition and factored out common parts in the definition
typo

%Math notation
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\longrightarrow}}
\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}

@@

%Functions
\newcommand{\plainfun}[1]{\mathit{#1}}
\newcommand{\decode}{\plainfun{decode}}
\newcommand{\encode}{\plainfun{encode}}
\newcommand{\encodePerm}{\plainfun{encodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}
\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc(#1)}}

@@

@@

                   & \{ \npair{\stdcap[\readonly]} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
                   & \{ \npair{\readwrite} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
\union \\ & \dots\text{Continues on next page.}
                   & \{ \npair{\stdcap[\exec] = c} \mid \\
                   & \quad (\exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
                   & \quad \land \\
                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}

@@

                   & \{ \npair{\stdcap[\entry]} \mid \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\
                   & \quad \land \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\}

@@

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
                           & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
                           & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
                           & \qquad \Rightarrow \npair[n-i]{\heap(0)} \in \stdvr(W) \land \\
                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap(0)} \in W(r)(W) )\}
\end{align*}
The ``good observations'' are as follows
\begin{itemize}
\item The execution does not stop in $n$ steps. 
\item The execution stops with a $\failed$.
\item The execution halts in at most $n$ steps with a valid value in in address 0 and the value at address 0 is also valid with respect to the world.
\end{itemize}
Address 0 is the designated I/O address.

@@

\end{definition}
\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  If for all $n \in \nats$, $W \in \Worlds$, and register files $\reg$, $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
\end{lemma}

>>>>

%Math notation
\newcommand{\restrictfun}[1]{|_{#1}}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\longrightarrow}}
\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}

@@

%Functions
\newcommand{\plainfun}[1]{\mathit{#1}}
\newcommand{\decode}{\plainfun{decode}}
\newcommand{\encode}{\plainfun{encode}}
\newcommand{\encodePerm}{\plainfun{encodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}

\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}
\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc(#1)}}

\newcommand{\readCond}[1]{\plainfun{readCondition}(#1)}
\newcommand{\writeCond}[1]{\plainfun{writeCondition}(#1)}
\newcommand{\execCond}[1]{\plainfun{executeCondition}(#1)}
\newcommand{\entryCond}[1]{\plainfun{entryCondition}(#1)}

@@

\[
  \iota_{\var{start},\var{end}} \; : \; \Worlds \monfun \UPred{\Addrs \parfun \Heaps}
\]
\begin{align*}
  \iota_{\var{start},\var{end}}(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                     & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
\end{align*}

\begin{align*}
  \readCond{n,W,\start,\addrend} = & \;\exists r \in \RegionName \ldotp \\
                                   & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
  \writeCond{n,W,\start,\addrend} = & \; \exists r \in \RegionName \ldotp \\
                                   & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\
  \execCond{n,W,\start,\addrend} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                   & \; \quad \forall \npair[n']{v_1},\dots,\npair[n']{v_k} \in \stdvr(W'),\addr' \in [\start, \addrend] \ldotp \\
                                   & \; \qquad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \\
                                   & \; \qquad \quad \in \stder(W') \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                   & \; \quad \forall \npair[n']{v_1},\dots,\npair[n']{v_k} \in \stdvr(W') \ldotp \\
                                   & \; \qquad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \\
                                   & \; \qquad \quad \in \stder(W') \\
\end{align*}

@@

                   & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
%                   & \quad \exists r \in \RegionName \ldotp \\
%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
                   & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
%                   & \quad \exists r \in \RegionName \ldotp \\
%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
%                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
%                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
%                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
%                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
                   & \{ \npair{\stdcap[\exec]} \mid \\
                   & \quad\readCond{n,W,\start,\addrend} \land \\
%                   & \quad (\exists r \in \RegionName \ldotp \\
%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
                   & \quad \execCond{n,W,\start,\addrend} \}
%                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
%                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}

@@

                   & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
%                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
%                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

                   & \quad \writeCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend} \}
%                   & \quad \exists r \in \RegionName \ldotp \\
%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
%                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
%                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
%                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
%                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\
%                   & \quad \land \\
%                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
%                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}

@@

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
                           & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
                           & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
                           & \qquad \Rightarrow \npair[n-i]{\heap(0)} \in \stdvr(W) \land \\
                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap \restrictfun{\{0\}}} \in W(r)(W) )\}
\end{align*}
The ``good observations'' are as follows
\begin{itemize}
\item The execution does not stop in $n$ steps. 
\item The execution stops with a $\failed$.
\item The execution halts in at most $n$ steps with a valid value in in address 0 and the value at address 0 is also valid with respect to the world.
\end{itemize}
Address 0 is the designated I/O address.

@@

\end{definition}
\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  For any $n \in \nats$, $W \in \Worlds$, and register file $\reg$, if  $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
\end{lemma}
